---
title: "Mario Kart Data Visualization"
author: "Vivian Guetler"
output: html_notebook
---

The data for this week's TidyTuesday comes from Mario Kart World Records and contains world records for the classic racing game on the Nintendo 64.

Data source: <https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-05-25/readme.md>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries
```{r}
library(tidyverse)
library(lubridate)
library(patchwork)
library(tidytext)
library(tilemaps)
library(showtext)
library(skimr)

# add fonts
font_add_google("Odibee Sans", "odibee-sans")
font_add_google("Rubik", "rubik")
showtext_auto()
```

# Load data
```{r}
records <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-25/records.csv')
drivers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-25/drivers.csv')

```

# View the data
```{r}
glimpse(drivers)
glimpse(records)
```

# EDA 
```{r}
skim(drivers)
```

```{r}
skim(records)
```

# Drivers
```{r}
drivers %>%
  count(player, sort = T)
```

# Player's nationality - USA leads, followed by Australia and Canada
```{r}
drivers %>%
  count(nation, sort = T)
```

# Tracks
```{r}
records %>%
  count(track, sort = T)
```
# Plots

# First Merge the two datasets
```{r}
records_drivers <- records %>%
  left_join(drivers, by="player") %>%
  group_by(year) 
  
```

# plot records over time
```{r}
records_drivers %>%
  summarise(records = mean(records,na.rm=TRUE)) %>% 
  ggplot(aes(x = year, y = records)) +
  geom_line(color="blue") + 
  geom_point(size = 1.1) +
  ggthemes::theme_hc(bgcolor = "darkunica") +
  labs(y = "Average Records",
       title = "Average Records Over Time")
```

# Which track is the fastest?

```{r}
records_drivers %>%
  group_by(track) %>%
  summarise(time = median(time)) %>%
  ggplot(aes(reorder(x = track,time), time)) + 
  geom_col(fill = "blue") +
  coord_flip() +
  labs(y = "Median Time(s)",
       x = "",
       title = "Median Time per Track") +
  theme(legend.position = "bottom",
        plot.title = element_text(family = "odibee-sans", size = 16),
        plot.subtitle = element_text(),
        plot.caption = element_text(size = 8, color = "grey50"),
        axis.ticks = element_blank(),
        legend.margin=margin(t = -0.5, unit='cm'))
```

# remove the missing 
```{r}
records_drivers <- filter(records_drivers, !is.na(nation))
```

```{r}
records_drivers$count <- 1
rd2 <- records_drivers %>% 
  group_by(track, nation) %>% 
  summarise(all_records = sum(count))
```

# heat map of tracks and records by country

```{r}
p1 <- ggplot(data=rd2, aes(x=nation, y=track, fill=all_records)) + 
  geom_tile() +
  scale_fill_gradient("Number of\nWorld Records\n", 
                      low="#feb24c", high="#800026") +
  labs(subtitle="Mario Kart 61: The number of world records by laps, \n track and players in different countries.\n") +
  theme(panel.background = element_rect(fill = "#cce6ff"),
        plot.background = element_rect(fill = "#cce6ff"),
        legend.background = element_rect(fill = "#cce6ff"),
        plot.title = element_text(colour = "#800026", size=20, face="bold", hjust = 0, family="odibee-sans"),
        plot.subtitle = element_text(colour = "#800026", size=10, hjust = 0, family="rubik"),
        legend.position="right",
        plot.margin = unit(c(0.5, 0.3, 0.3, 0.3), "cm"), #top, right, bottom, left
        legend.spacing.x = unit(0.5,"cm"),
        legend.title = element_text(colour = "#800026", size=10, hjust = 0.5, family="odibee-sans"),
        legend.text = element_text(colour="#800026", size=10, family="rubik", hjust = 0.5),
        axis.title.x= element_blank(),
        axis.title.y= element_blank(),
        axis.text.x=element_text(colour = "#800026", size=10, 
                                 hjust = 0, vjust=0.5, family="rubik", angle=-90),
        axis.text.y=element_text(colour = "#800026", size=10, 
                                 hjust = 1, family="rubik"),
        axis.ticks.x=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()
  )

                    
```

# top 10 players
```{r}
p2 <-drivers %>% 
  group_by(player) %>%
  summarise(record_count = sum(records,na.rm = TRUE)) %>% 
  arrange(desc(record_count)) %>% 
  head(10) %>%
  ggplot(aes(reorder(player,record_count),record_count, fill=player)) +
  geom_col() +
  coord_flip() +
  labs(x = "",
       y = "Record Count",
       subtitle = "Top 10 players\nwith the most world records\n") +
  theme(panel.background = element_rect(fill = "#cce6ff"),
        plot.background = element_rect(fill = "#cce6ff"),
        plot.title = element_text(colour = "#800026", size=20, face="bold", hjust = 0, family="odibee-sans"),
        plot.subtitle = element_text(colour = "#800026", size=10, hjust = 0, family="rubik"),
        legend.position="none",
        plot.margin = unit(c(0.5, 0.3, 0.3, 0.3), "cm"), #top, right, bottom, left
        axis.title.x= element_blank(),
        axis.title.y= element_blank(),
        axis.text.x=element_text(colour = "#800026", size=10, 
                                 hjust = 0, vjust=0, family="rubik"),
        axis.text.y=element_text(colour = "#800026", size=10, 
                                 hjust = 0, family="rubik"),
        axis.ticks.x=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()
  )
```


# Combine plots

```{r}
plots <- p1 + p2 + plot_layout(ncol = 2) +
  plot_annotation(
    caption = '@vivfiona | TidyTuesday | Data: Mario Kart World Records',
    title = '\nMARIO KART WORLD RECORDS') &
  theme(panel.background = element_rect(fill = "#e5e7e9", colour="#e5e7e9"),
        plot.background = element_rect(fill = "#e5e7e9", colour="#e5e7e9"),
        plot.title = element_text(colour = "#800026", size=20, face="bold", hjust = 0.5, family="odibee-sans"),
        plot.caption = element_text(colour = "#800026", size=10, hjust = 1, family="rubik"))


```

# Save the plots

```{r}
ggsave (
  width = 9,
  height = 4.5,
  filename = "mariokart.png",
  device = "png",
  dpi = 300)
```


