---
title: "Internet Access West Virginia"
author: "Vivian Guetler"
date: "5/11/2021"
output: html_document
---

### Internet Access in the US
# Data source: Microsoft
# Tidytuesday : <https://github.com/rfordatascience/tidytuesday/tree/master/data/2021/2021-05-11>

# Codes inspired by and adapted from: Jesse Mostipak <https://twitter.com/kierisi/status/1391915889331683331> and @leeolney3


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup - Load the packages needed 
```{r}
library(tidyverse) #for plotting 
install.packages("ggridges") #for ridge plots
library(ggridges)
library(skimr) # for data analysis
#install.packages("janitor") #for cleaning column names
library(janitor)
install.packages("patchwork") # for combining separate ggplots into same graphic
library(patchwork)
#install.packages("wesanderson") # for color palettes
library(wesanderson)
library(sf) # for maps
library(rnaturalearth)
library(rnaturalearthdata)
#install.packages("urbnmapr") # for maps
#library(urbnmapr)
library(showtext)
## Loading Google fonts (https://fonts.google.com/)
font_add_google("Ubuntu", "buntu")
font_add_google("Oxygen", "oxy")
## Automatically use showtext to render text
showtext_auto()
```

# Load the data
```{r}
broadband <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-11/broadband.csv')

broadband_zip <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-11/broadband_zip.csv")
```

# Glimpse the data

```{r}
glimpse(broadband)
```
```{r}
glimpse(broadband_zip)
```
```{r}
skim(broadband) #eda/descriptive stats
```

```{r}
skim(broadband_zip)
```

# Standardizing column headers - right now all caps

```{r}
broadband_new <- broadband %>%
  janitor::clean_names()
```

```{r}
names(broadband_new) # worked, now small letters
```

# Do the same for the second dataset
```{r}
broadband_zip <- broadband_zip %>%
  janitor::clean_names() #now column names are all small letters
```

# Analysis
# What is the distribution of broadband usage by state

# First convert broadband_usage to numeric

```{r}
broadband_new %>%
  mutate(broadband_usage = as.numeric(broadband_usage))
```

# NA's introduced by coercion.... look closer

```{r}
broadband_new %>%
  View() # some states have a dash (missing)
```

# Remove the dash, convert broadband_usage as numeric

```{r}
broadband_plot <- broadband_new %>%
  filter(broadband_usage != "-") %>%
  mutate(broadband_usage = as.numeric(broadband_usage), broadband_availability_per_fcc = as.numeric(broadband_availability_per_fcc))
```


#check avg broadband availability by state
```{r}
colnames(broadband_new)
st_broadband <- broadband_new %>% group_by(st)
st_broadband <- st_broadband %>% summarise(avg = mean(as.numeric(broadband_availability_per_fcc), na.rm=T))

max_state <- st_broadband$st[which.max(st_broadband$avg)] #max state = CT
min_state <- st_broadband$st[which.min(st_broadband$avg)] #min state = AR
```

# Boxplot of broadband_usage by state

```{r}
broadband_plot %>%
  ggplot(aes(x = st, y = broadband_usage)) +
  geom_boxplot()
```

# Reorder the plot by median

```{r}
broadband_plot %>%
  ggplot(aes(x = reorder(st, broadband_usage), y = broadband_usage)) + geom_boxplot()
```

# make the boxplot easier to read

```{r}
broadband_plot %>%
  ggplot(aes(x = reorder(st, broadband_usage), y = broadband_usage)) + 
  geom_boxplot() +
  theme_bw() +
  coord_flip()

```

# Try another plot ; ridge plot

```{r}
broadband_plot %>%
  ggplot(aes(x = broadband_usage, y = reorder(st,  broadband_usage))) +
  geom_density_ridges() +
  labs(
    title = "Internet Usage in the US",
    subtitle = "Percent of people per county that use the internet at broadband speed ",
    caption = "Source: Microsoft/The Verge")
```

# Save ridge plot

```{r}
ggsave ("broadband_ridgeplot", last_plot(), 
        device = png)
```


# other plots

```{r}
plot1 <- broadband_plot %>%
  ggplot(aes(x = broadband_usage, y = reorder(st,  broadband_usage), fill = stat(x))) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "broadband use", option = "C") +
  labs(
    title = "Internet Usage in the US",
    subtitle = "Percent of people per county that use the internet at broadband speed ",
    caption = "@Vivfiona, Data Source: Microsoft & The Verge",
    x = "Broadband Usage",
    y = "State") +
  theme_minimal(base_size = 10) + 
  theme(axis.text.y = element_text(vjust = 0))
  #theme_ridges(grid = FALSE, center_axis_labels = TRUE) # removes the grid
```

# Save the plot

```{r}
ggsave (
  width = 8,
  height = 4.5,
  filename = "broadband_ridgeplot.png",
  device = "png")
```



```{r}
broadband_plot %>%
  ggplot(aes(y = reorder(st,  broadband_usage))) +
  geom_bar() +
  labs(
    title = "Internet Usage in the US",
    subtitle = "Percent of people per county that use the internet at broadband speed ",
    caption = "Source: Microsoft/The Verge")
```

# Maps
```{r}
install.packages("devtools")
devtools::install_github("UrbanInstitute/urbnmapr")
library(urbnmapr)
```

# Get counties map - focus on WV
```{r}
counties_wv <- get_urbn_map("counties", sf=TRUE)
```

```{r}
glimpse(broadband_plot)
```

# plot maps of the broadband availability in WV
```{r}
pl1 <- counties_wv %>%
  filter(state_abbv == "WV" ) %>%
  left_join(broadband_plot, by = "county_name") %>%
  ggplot() +
  geom_sf(mapping = aes(fill = broadband_availability_per_fcc), color = "#ffffff", size = 0.25) +
  coord_sf(datum = NA) + 
  scale_fill_gradientn(labels = scales::percent,
                       colours = wes_palette("Darjeeling1", 100, type = "continuous"),
                       na.value = "grey",
                       trans = "reverse") +
  theme_minimal(base_size = 10 ) +
  theme(plot.subtitle = element_text(size = 8, family = "oxy", hjust = 0.5),
        plot.title = element_text(size = 14, family = "buntu", face = "bold")) +
  guides(fill = guide_colorbar(title.position = "top",
                               title.hjust = .5,
                               reverse = TRUE,
                               barheight = unit(10, "lines"),
                               barwidth = unit(.5, "lines"))) +
  labs(fill ="", title = "Broadband Availability", 
       subtitle = str_wrap("% of people per county with access to fixed terrestrial broadband at speeds of 25 Mbps/3 Mbps as of the end of 2017"))
  
```

# plot map of the broadband usage in WV

```{r}
pl2 <- counties_wv %>%
  filter(state_abbv == "WV" ) %>%
  left_join(broadband_plot, by = "county_name") %>%
  ggplot() +
  geom_sf(mapping = aes(fill = broadband_usage), color = "#ffffff", size = 0.25) +
  coord_sf(datum = NA) + 
  scale_fill_gradientn(labels = scales::percent,
                       colours = wes_palette("Darjeeling1", 100, type = "continuous"),
                       na.value = "grey",
                       trans = "reverse") +
  theme_minimal(base_size = 10 ) +
  theme(plot.subtitle = element_text(size = 8, family = "oxy", hjust = 0.5),
        plot.title = element_text(size = 14, family = "buntu", face = "bold")) +
  guides(fill = guide_colorbar(title.position = "top",
                               title.hjust = .5,
                               reverse = TRUE,
                               barheight = unit(10, "lines"),
                               barwidth = unit(.5, "lines"))) +
  labs(fill ="", title = "Broadband Usage", 
       subtitle = str_wrap("Percent of people per county that use the internet at broadband speeds", 40))

```

# Combine the 2 plots

```{r}
pl1 + pl2 + plot_annotation(title = "Internet Access in West Virginia", caption = "#TidyTuesday | @Vivfiona | Data from Microsoft & The Verge") & theme(plot.title = element_text(hjust = 0.5, size = 20, family = "oxy", face = "bold"), plot.caption = element_text(hjust = 0.5))
```

# save the plots

```{r}
#ggsave ("broadband_maps",width = 7, height = 5, dpi = 300)

ggsave (
  width = 9,
  height = 4.5,
  filename = "broadband_maps.png",
  device = "png",
  dpi = 300)
```

